Technical Report

### 2018-01-12 fully-abstract-comp/technical-report.tex
Wrote up suggestion for updated LR. Removed K-relation and wrote new cases for return capabilities (Future world relation still to be updated.)

Split cannot create ``empty capabilities'' (a capability that governs no segment of the memory, i.e.\ a capability where the base address is greater than the end address). We partly do not allow this out of convenience as it makes the implementation of call simpler. We do not need empty capabilities as they have no semantic value in the sense that they allow you to do no more than a piece of data.

@@

@@

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{c,S},k_{d,S}}{k_{c,T},k_{d,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rretd,\rretc,\rdata\})(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

\[
  \begin{array}{rl}
  \lrk(W) = &\left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
    \begin{array}{l}
      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
      \npair[n]{[\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrd(\sigma,\_)}{k_{d,T}}]} \in \lrp(W) \wedge \\
      \forall n' \leq n, W' \pubft W \ldotp \\
      \quad \forall W_\lrr, W_\lrm \ldotp  \\
      \qquad\memSat[n']{\stpair[.]{\ms_S,(\opc,\ms_{\stk,f}) :: \stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\opc = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c+12) \wedge\\
      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
      \qquad\npair[n']{\left(\array{l}\src{\reg_S\update{\rstk}{\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}}},\\
                               {\reg_T\update{\rstk}{((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)}}\endarray \right)}\in \lrr(\{r_\stk\})(W_\lrr) \wedge\\ 
      \qquad\npair[n']{[(\opc,\opc)]} \in \lrp(W')\\
      \qquad\Rightarrow\\
      \qquad \npair[n']{\left(\arraycolsep=0pt\array{l}
                          \left(\arraycolsep=0pt\array{l}
                            \src{\ms_S,\reg_S\update{\pcreg,\rstk}{\opc,\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}},}\\
                            \src{\stk, \ms_\stk \uplus \ms_{\stk,f}}
                          \endarray \right) \\
                          \left(\ms_T,\reg_T\update{\pcreg,\rstk}{\opc,((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)} \right)
                        \endarray \right)} \in \lro(W'\oplus W_\lrr \oplus W_\lrm)
    \end{array}
  \right\}
% \\ \cup
%     & \left\{ \npair{\left( k_{c,S},k_{d,S},k_{c,T},k_{d,T} \right)} \middle|
%       \begin{array}{l}
%         \pi_\sigma(k_{c,S}) = \pi_\sigma(k_{d,S}) \wedge \pi_\sigma(k_{c,T}) = \pi_\sigma(k_{d,T}) \wedge \npair{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \\
%         \forall n' \leq n, W' \pubft W \ldotp \\
%         \quad \npair[n']{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \Rightarrow\\
%         \qquad \forall \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W'_\lrr) \ldotp 
%                \forall \memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W'_\lrm} \ldotp\\
%         \qquad \quad \npair[n']{\left(
%         \arraycolsep=0pt\array{l}
%           (\src{\ms_S},\src{\reg_S}\update{\pcreg,\rdata}{\pi_\vsc(\src{k_{c,S}}),\pi_\vsc(\src{k_{d,S}})},\src{\stk},\src{\ms_\stk}),\\
%           (\ms_T,\reg_T\update{\pcreg,\rdata}{\pi_\vsc(k_{c,T}),\pi_\vsc(k_{d,T})})
%         \endarray\right)} \in \lro(W' \oplus W'_\lrr \oplus W'_\lrm)
%       \end{array}
%      \right\}
  \end{array}
\]

@@

\lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
\[
  \lrv(H_\sigma,W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
    \begin{array}{l}
      \vsc_C = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge \\
      \forall (\src{c},c') \in H_\sigma(\sigma) \ldotp \\
      \quad \exists r \in \addressable{\linear,W} \ldotp \\
      \qquad  \exists \stpair{\ms}{\ms} \ldotp \\
      \qquad \quad \dom(\src{\ms_S}) = \dom(\ms_T) = [\pi_\baddr(c), \pi_\eaddr (c)] \wedge \stareg{\spao} \nequal W(r) \wedge \\
      \qquad \qquad \forall W' \privft W \ldotp \\
      \qquad \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
      \qquad \qquad \qquad { \color{red} \npair{(\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)}, \src{c}, \sealed{\sigma,\vsc_c}, c')}} \in \lrk(W')
    \end{array}
    \right\}\cup\\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \;\middle| \;
    \begin{array}{l}
      \vsc_D = ((\rw,\linear),\baddr,\eaddr,\baddr-1) \wedge\\
      \exists r \in \addressable{\linear,W} \ldotp \\
      \quad  \exists \stpair{\ms}{\ms} \ldotp \\
      \qquad \dom(\src{\ms_S}) = \dom(\ms_T) = [\baddr, \eaddr] \wedge \stareg{\spao} \nequal W(r) \wedge \\
      \qquad \forall W' \privft W \ldotp \\
      \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
      \qquad \qquad { \color{red} \forall (\src{c}, c') \in H_\sigma (\sigma) \ldotp \npair{\src{c,\retptrd(\sigma,\baddr,\eaddr)}, c' , \sealed{\sigma,\vsc_D}} \in \lrk(W') }%\text{\lau{This needs some work - specifically what i c and c' and do we need to seal them?}}

%    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
%      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
    \end{array}
    \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
      \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
      \quad \npair{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lre(W) \wedge \\
      \quad \npair{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lre(W)
    \end{array}
    \right\}\cup\\
%
    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    % \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{r l l}
      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{r l l }
      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
      \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
      \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
    \end{array}
    \right\}
  \end{array}
\]
\lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}

>>>>

Split cannot create ``empty capabilities'' (a capability that governs no segment of the memory, i.e.\ a capability where the base address is greater than the end address). We partly do not allow this out of convenience as it makes the implementation of call simpler. We do not need empty capabilities as they have no semantic value in the sense that they allow you to do essentially the same as a piece of data.

@@

\dominique{4-12-2017: $e_\stk$ is no longer stored in $\ms_{\stk,\priv}$ in updated call implementation.}
\lau{12-12-2017: Correct. Please feel free to correct these things.}

@@

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

@@

\lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
\[
  \lrv(H_\sigma,W) =
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
    \begin{array}{l}
      (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
      \npair{(\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})} \in \lrp(W)
    \right\}\cup\\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
      \forall n' \leq n \ldotp \forall W' \privft W \ldotp \\
      \quad \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
      \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}{W_\lrm}} \wedge \\
      \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata,\rstk,\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
      \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
      \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
      \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
      \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
      \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
      \quad W_\var{new} = (W' \uplus W_\lrr \uplus W_\lrm \uplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
      \qquad \Rightarrow \\
      \qquad \quad \npair[n']{\left(\array{l}(\ms_S,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0},\stk,\ms_\stk \uplus \ms_f),\\
                                             (\ms_T,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?)})
                                    \endarray\right)} \in \lro(W_\var{new})
    \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
      \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
      \quad \npair{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lre(W) \wedge \\
      \quad \npair{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lre(W)
    \right\}\cup\\
%
    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    % \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{r l l}
      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{r l l }
      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
      \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
      \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
    \right\}
\]
where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.
\lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
